---
title: "Crime_data_US"
author: "kRushnAM"
date: "2026-01-11"
output: html_document
---

* We started with a data set downloaded from https://catalog.data.gov/dataset/crime-data-from-2020-to-present

[at first I thought  to clean the data-set and query it using MySQL (I clearly over estimated my skills) and the Unicode col names made me loose my sleep]

so ultimately I opened my trusty R studio and installed packages for Data cleaning, EDA and Viz.

```{r}
View(crime_data_details)
```

As we can see the complete_rate for cols like crime_cd 4,3,2, cross_street,weapon-used_cd, weapon_desc, mocodes is less, I'm creating a new table with rest of the cols for the purpose of our analysis

after dropping those cols and removing NA values, we got a clean data set called "crimes_data" (clean crimes! irony!!!)

```{r}
View(crimes_data)
```

Initially taught to split the date_occured and date_reported into date, month, year but that leads to more cols and we aren't gonna do a prediction analysis so why to bother!

so, the time_occ is in text format, but we need it in HH:MM format. Hence, lets change that into its original format.

I used str_pad and then sub functions to change the time_occured into
1.A 4 character text
2.Then sub function converted it into HH:MM format.
Easy! Peasy!!!

Now, when I looked at Vict_age I saw -4,-2 age. That is either a typo or we have some really scary ghosts out there irritating baby ghosts.
for the sanity of this analysis lets 

Oops! while doing all this, I accidentally closed R studio. That's why, they say we need to save the DF first. Hence, Crimes_data has been saved as .csv file [I'm gonna load that csv into MySQL workbench and gonna torture data with illogical queries]

We have 
1.cleaned col names
2. removed NA vals
3. Morphed the dates from text to date
4. time from text to time
5. removed -ve values from vict_age
so the only thing left is Visualization [Drum_role]

Before jumping into the ggplot, I had to
1.Unpack the date col into date, month and year (seriously! How did I forgot)
2. for first viz we gonna concentrate on years, months and total crimes reported